---
title: "Application to Ozone Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ozone}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(earth.dof.patch)
library(earth)
```

## Ozone Data

The Ozone data is shift with `earth` package, called `ozone1`

To check the performance in higher dimension, we can manually create 50 nuisance features.

```{r, eval=FALSE}
set.seed(1234)
ozone1 = cbind(ozone1, matrix(rnorm(nrow(ozone1) * 50), ncol = 50))
```

```{r}
head(ozone1)
```

```{r}
mod = earth(O3 ~ ., data = ozone1)
```

```{r}
summary(mod)
```

then calculate the corrected penalty given the fitted earth object `mod`,

```{r}
res = correct_df(mod)
```

refit the earth model with new penalty

```{r}
mod1 = earth(O3 ~ ., data = ozone1, penalty = res$penalty)
```

```{r}
summary(mod1)
```


perform a 5-fold cross-validation

```{r}
cv_compare = function(nfold = 5, degree = 2) {
  n = nrow(ozone1)
  folds = rep(1:nfold, each = n / nfold)
  folds = folds[sample(1:n, n)] # re-order
  errs = matrix(0, nrow = 2, ncol = nfold)
  corrected_penalty = NULL
  for (k in 1:nfold) {
    train = ozone1[folds != k, ]
    test = ozone1[folds == k, ]
    mod = earth(O3 ~ ., data = train, degree = degree)
    if (is.null(corrected_penalty)) {
      dfs = correct_df(mod)
      corrected_penalty = dfs$penalty
      cat("corrected penalty = ", dfs$penalty, "\n")
    }
    mod1 = earth(O3 ~ ., data = train, degree = degree, penalty = corrected_penalty)
    pred = predict(mod, newdata = test)
    pred1 = predict(mod1, newdata = test)
    errs[, k] = c(mean((test$O3 - pred)^2), mean((test$O3 - pred1)^2))
  }
  mu = rowMeans(errs)
  se = apply(errs, 1, sd) / sqrt(nfold)
  cat("CV error = \n")
  print(mu)
  cat("CV error se = \n")
  print(se)
  return(errs)
}
```

when we only use linear terms,

```{r}
set.seed(1234)
errs.d1 = cv_compare(5, 1)
```

and if we allow second-degree interactions,

```{r}
set.seed(1234)
errs.d2 = cv_compare(5, 2)
```
