---
title: "Application to the Spam Email Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{spam}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(earth)
library(earth.dof.patch)
```

The `spam` data has been frequently used in The Elements of Statistical Learning (ESL) book. The data consists of information from 4601 email messages, in a study to try to predict whether the email was spam (i.e., junk email). The data can be accessible from <https://archive.ics.uci.edu/dataset/94/spambase> or <https://github.com/szcf-weiya/ESL-CN/tree/master/data/Spam>.

Here, we also include the data into our package, and one can simply load it as follows:

```{r}
data("spam")
head(spam)
data("flag.esl.trainset")
spam.train = spam[flag.esl.trainset == 0, ]
spam.test = spam[flag.esl.trainset == 1, ]
```

where `flag.esl.trainset` is the indicator for training set and test set used in the ESL site.

Now we compare the performance of the original MARS (via `earth`) and our proposed MARS with corrected degrees of freedom.

```{r}
compare_class_table = function(degree = 1) {
  mod = earth(class ~ ., data = spam.train, degree = degree)
  pred = predict(mod, newdata = spam.test, type = "class")
  tbl = table(pred, truth = spam.test$class)
  cat("the confusion table is:\n")
  print(tbl)
  cat("the confusion table in proportion is:\n")
  print(tbl / sum(tbl))
  cat("the error rate is: ", (tbl[2] + tbl[3]) / sum(tbl), "\n" )
  cat("==============\n")
  dfs = correct_df(mod)
  cat("use corrected penalty = ", dfs$penalty, "\n")
  mod1 = earth(class ~ ., data = spam.train, degree = degree, penalty = dfs$penalty)
  pred1 = predict(mod1, newdata = spam.test, type = "class")
  tbl1 = table(pred1, truth = spam.test$class)
  cat("the confusion table is:\n")
  print(tbl1)
  cat("the confusion table in proportion is:\n")
  print(tbl1 / sum(tbl1))
  cat("the error rate is: ", (tbl1[2] + tbl1[3]) / sum(tbl1), "\n" )
}
```

Firstly, when we only use linear terms, the performance are

```{r}
compare_class_table(1)
```

Furthermore, when we allow second-degree interactions, the performance are

```{r}
compare_class_table(2)
```
